\section{Connectivity}
\subsection[k-connectivity]{$k$-connectivity}

\begin{definition}
    The \emph{connectivity number}\index{connectivity number} 
    $\kappa(G)$ is the minimum number of vertices in $S \subseteq 
    V(G)$ such that $G - S$ is disconnected or contains only 
    one vertex.
\end{definition}

\begin{remark}
    The latter conditions handles the case of complete graphs.
\end{remark}

\begin{definition}
    A graph $G$ is $k$-connected\index{k@$k$-connectivity} if $\kappa(G) \ge k$.
\end{definition}

Alternatively: $G$ is $k$-connected if the removal of at most $k-1$ vertices always
results in a connected graph with at most two vertices.

\begin{proposition}
    \begin{enumerate}
        \item $\kappa(G) \le \delta(G)$
        \item $\kappa(G) \le \beta(G)$
        \item $\kappa(G) \le n(G) - 2$
    \end{enumerate}
\end{proposition}

%\begin{proof}
%    % TO DO
%\end{proof}

\begin{theorem}
    The minimum number of edges in a $k$-connected graph of 
    order $n > k > 2$ is $\ceil{nk/2}$.
\end{theorem}

\begin{proof}
    Using \emph{Harary graphs}\index{Harary graph} $H_{n,k}$.
    % TO DO
\end{proof}

\datum{2024-10-24}

\begin{definition}
    A set $F \subseteq E(G)$ is a \emph{disconnecting set}\index{disconnecting set}
    if $E - F$ is disconnected
\end{definition}

\begin{definition}
    Let $\emptyset \neq A \subsetneq V(G)$ and let $E(A, \overline{A})$ be 
    the set oe edges between $A$ and $\overline{A}$. Then 
    $E(A, \overline{A})$ is an \emph{edge cut}.\index{edge cut}
\end{definition}

Clearly, an edge cut is a disconnecting set. Similarly, 
a minimal disconnecting set is an edge cut.

\begin{definition}
    The \emph{edge-connectivity numver} of $G$, 
    $\kappa'(G)$ is the minimum number of edges in a 
    disconnecting set.
\end{definition}

\begin{definition}
    A graph is \emph{$k$-edge-connected}\index{k@$k$-edge-connected}
    if the removal of less than $k$ edges always leaves a connected graph.
\end{definition}

Equivalently, a graph $G$ is $k$-edge-connected if $k \le \kappa'(G)$.

\begin{theorem}
    Let $G$ be a simple graph with $n(G) \ge 2$. Then 
    $\kappa(G) \le \kappa'(G) \le \delta(G)$.
\end{theorem}

%\begin{proof}
    % TO DO
%\end{proof}

\begin{corollary} Let $G$ be a graph.
    \begin{itemize}
        \item If $G$ is $k$-connected, then $G$ is 
        $k$-edge-connected.
        \item The minimum number of edges in a $k$-edge-connected
        graph on $n > k \ge 2$ vertices is $\ceil{kn/2}$.
    \end{itemize}
\end{corollary}

%\begin{proof}
    % TO DO
%\end{proof}

\begin{theorem}[Whitney]\index{Whitney's Theorem}
    A graph $G$ is $2$-connected if and only if for every 
    pair of distinct vertices $u,v \in V(G)$ there 
    exist two internally disjoint $u,v$-paths.
\end{theorem}

\begin{lemma}[Expansion]\index{Expansion Lemma}
    If $G$ is a $k$-connected and we add a new vertex $v$ 
    and $k$ incident edges to the graph, we obtain a 
    $k$-connected graph.
\end{lemma}

%\begin{proof}
    % TO DO
%\end{proof}

\datum{2024-11-07}

\begin{theorem}
    If $G$ is a graph with $n(G) \ge 3$, then the following statements are equivalent.
    \begin{enumerate}[(a)]
        \item $G$ is $2$-connected.
        \item $G$ is connected and there is no cut-vertex.
        \item For every pair of distinct vertices $u,v \in V(G)$ there 
        exist two internally disjoint $u,v$-paths.
        \item Every pair of distinct vertices lies on a common cycle.
        \item Every pair of edges lies on a common cycle and $\gamma(G) \ge 1$.
    \end{enumerate}
\end{theorem}

\begin{lemma}[Subdivision]\index{Subdivision Lemma}
    Let $G'$ be a graph obtained from $G$ by subdividing an edge. Then $G$ 
    is $2$-connected if and only if $G'$ is $2$-connected
\end{lemma}

\begin{definition}
    In a graph $G$, a path $P$ is an \emph{open ear}\index{open ear}\index{ear} if all internal
    vertices of $P$ are of degree $2$ and the end vertices are of degree at least $3$.
\end{definition}

\begin{definition}
    An \emph{open ear decomposition} of $G'$ is a sequence $P_0, P_1, \dots, P_k$ such 
    that $P_0$ is a cycle and for every $i \in [k]$ $P_i$ is an open ear in 
    the graph $G_i = P_0 \cup P_1 \cup \cdots \cup P_i$ and $G_k = G$.
\end{definition}

\begin{theorem}
    A graph $G$ is $2$-connected if and only if it has an open ear decomposition.
\end{theorem}

\begin{proposition}
    A graph $G$ is $2$-edge-connected if and only if it is connected and every edge of $G$ lies in a cycle.
\end{proposition}

\begin{definition}
    A \emph{closed ear}\index{closed ear} in $G$ is such a cycle that only one vertex has degree at least $4$.
\end{definition}

\begin{definition}
    A \emph{closed ear decomposition}\index{closed ear decomposition} of $G$ is $P_0, P_1, \dots, P_k$ such 
    that $P_0$ is a cycle and for every $i \in [k]$ $P_i$ is an open/closed ear in the graph $G_i = P_0 \cup P_1 \cup \dots \cup P_k$.
\end{definition}

\begin{theorem}
    A graph $G$ is $2$-edge-connected if and only if it has a closed ear decomposition.
\end{theorem}

%\datum{2024-11-14}