\section{Exercises}
\subsection*{Exercise sheet 1}
\datum{2024-10-10}
\begin{enumerate}
    \item For each of the following graphs $G$, determine $\alpha(G)$, $\alpha'(G)$, $\beta(G)$, 
    and $\beta'(G)$:
    \begin{enumerate}
        \item $G = P_n$
        \item $G = C_n$
    \end{enumerate}
    
    \item Prove that $\frac{\beta(G)}{2} \leq \alpha'(G) \leq \beta(G)$ for any graph $G$.
    
    \item (From lectures) Let $G$ be a bipartite graph with bipartition $\{A, B\}$. The deficiency 
    of the subset $U \subseteq A$ is defined as:
    \[
    \defi_G(U) := \abs{U} - \abs{N_G(U)},
    \]
    and the deficiency of $G$ is defined as:
    \[
    \defi(G) := \max_{U \subseteq A} \defi_G(U).
    \]
    Prove that $\alpha'(G) = \abs{A} - \defi(G)$.
    
    \item Let $F = \set{F_1, \dots, F_n}$ be a family of subsets of a set $Y$. 
    Prove there are distinct elements $a_1, \dots, a_n$ such that $a_i \in F_i$ if and only if 
    $\abs{F} \leq \abs{\cup_{S \in F} S}$ for all $F \subseteq S$. 
    (Such a set is called a system of distinct representatives for $F$.)
    
    \item (Extra) Let $\Gamma$ be a finite group and $H \leq \Gamma$ with index $n$. 
    Let $L = \set{L_i}_{i=1}^n$ be the left cosets of $H$ and $R = \set{R_i}_{i=1}^n$ be the right cosets of $H$. 
    Prove that there is a subset $\set{h_1, \dots, h_n} \subseteq G$ such that $L = \set{h_i H}_{i=1}^n$ and 
    $R = \set{H h_i}_{i=1}^n$.
\end{enumerate}

\newpage
\subsection*{Exercise sheet 2}
\datum{2024-10-17}
\begin{enumerate}
    \item For each of the following graphs $G$, determine the number of maximum matchings:
    \begin{enumerate}
        \item $G = K_n$
        \item $G = K_{a,b}, a \leq b$
    \end{enumerate}
    
    \item Let $G$ be a graph such that $\delta(G) \geq 1$, with maximal matching $M$ and minimal edge cover $C$. 
    Prove the following equivalences:
    \begin{enumerate}
        \item $M$ is a maximum matching if and only if $M$ is contained in a minimum edge cover.
        \item $C$ is a minimum edge cover if and only if $C$ contains a maximum matching.
    \end{enumerate}
    
    \item Use deficiency (see sheet 1) to prove the König-Egerváry theorem. 
    (Hint: find a matching and a vertex cover of the same size using a subset of maximum deficiency).
    
    \item (From lectures) Let $G$ be a bipartite graph with bipartition $\set{A, B}$ such that $\abs{A} = \abs{B}$. 
    Prove that for $A$, Hall's condition holds if and only if Tutte's condition holds.
    
    \item Prove Theorem \ref{thm:BTF} (Berge-Tutte formula).
\end{enumerate}

\newpage
\section{Homework}

\begin{enumerate}
    \item Use König's Theorem to prove that every bipartite graph 
    $G$ has a matching of size $e(G)/\Delta(G)$. 
    Use this to show that any subgraph of $K_{n,n}$ with more 
    than $(k - 1)n$ edges has a matching of size at least $k$.

    \item Let $G$ be a graph with $\delta(G) \geq k$ and $n(G) 
    \geq 2k$. Prove that $\alpha'(G) \geq k$ 
    (Hint: you can use the Berge-Tutte formula).

    \item Let $G$ be an $r$-connected graph of even order, 
    and which contains no $K_{1,r+1}$ as an induced subgraph. 
    Prove that $G$ has a 1-factor.

    \item A trail in a graph $G$ is a list of vertices and 
    edges $v_0, e_1, v_1, e_2, \ldots, e_{k-1}, v_k$ such 
    that $e_i = v_{i-1}v_i$ and the edges $e_i$ are pairwise distinct.

    Let $G$ be a graph with $n(G) \geq 3$. Prove the following 
    are equivalent:
    \begin{enumerate}
        \item $G$ is 2-edge-connected.
        \item Every edge appears in a cycle.
        \item $G$ has a closed trail containing any specified 
        pair of edges.
        \item $G$ has a closed trail containing any specified 
        pair of vertices.
    \end{enumerate}

    \item Let $G$ be a simple graph. Prove that 
    $\chi(G) + \chi(G^c) \leq n(G) + 1$.

    \item A graph $G$ with $\chi(G) = k$ is called colour-critical, 
    or $k$-critical, if $\chi(H) < k$ for all proper subgraphs $H$ of $G$.

    Prove the following:
    \begin{enumerate}
        \item Every $k$-chromatic graph has a $k$-critical subgraph.
        \item A graph $G$ with no isolated vertices is colour-critical 
        if and only if $\chi(G - e) < \chi(G)$ for all $e \in E(G)$.
        \item Let $G$ and $H$ be $k$-critical graphs sharing exactly 
        $1$ vertex $v$, and with edges $vu \in E(G)$, $vw \in E(H)$. 
        Prove that $(G - uv) \cup (H - vw) \cup uw$ is $k$-critical.
        \item For all $k \geq 3$, use (c) to obtain a $k$-critical 
        graph other than $K_k$.
        \item For $n \geq 4$, $n \neq 5$, construct a 4-critical 
        graph with $n$ vertices.
    \end{enumerate}
\end{enumerate}